---
title: "Study 3"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prerequisites, message=FALSE, warning=FALSE, paged.print=FALSE, echo=FALSE}
rm(list = ls());
setwd("/Users/bsyiem/Documents/UnityProjects/PLMS/RScripts");

# for reading xlsx files
library("xlsx")

#used to easily plot qqplots and Anova
library("car");
library("ez");

#for two way Anova
library(tidyverse);
library(ggpubr);
library(rstatix);

#Multilevel Linear Model
library(lme4);
library(lmerTest);

#data formatting
library(dplyr);
library(tidyr);
library(data.table);

library(plyr);

# for equivalence test
library("TOSTER");

library(pwr);

library(rje);
```

```{r data-setup, include=FALSE}
mpath <- "../PythonScripts/Data/Participant_Data/"
mdata_file <- "data.xlsx"

my_data <- read.xlsx(paste(mpath,mdata_file, sep = ""), sheetIndex = 1)

my_data$pid <- as.factor(my_data$pid)
my_data$gender <- as.factor(my_data$gender)
my_data$mid_air <- as.factor(my_data$mid_air)
my_data$problem_solving <- as.factor(my_data$problem_solving)

my_data$q1 <- as.factor(my_data$q1)
my_data$q2 <- as.factor(my_data$q2)
my_data$q3 <- as.factor(my_data$q3)
my_data$q4 <- as.factor(my_data$q4)
my_data$q5 <- as.factor(my_data$q5)


#RATE CORRECT SCORE -> "a comparison of methods to combine speed and accuracy measures of performance: a rejoining on the binning procedure" - Andre Vandierendonck
#my_data$rcs <- sqrt(my_data$rcs) # since we have 0 values, we cannot use log transform and use a sqrt transform instead

#discard participants
to_discard <- c(5,7,25,29, 41,47,62)
my_data <- my_data[!(my_data$pid %in% to_discard),]

#outlier_discard <- c(2,30,39,11,28,34)
#my_data <- my_data[!(my_data$pid %in% outlier_discard),]
```

## Summaries

```{r summary, include=FALSE}
summary(my_data)
```

The following shows a summary of the data grouped by method (Physical, AR, RL_CLUSTER). Summary indicates very slight differences in mean completion times.

```{r completion-summary, echo=FALSE}
completionTime_summary <- my_data %>% group_by(method) %>% get_summary_stats(completion_time, type = "mean_sd")
completionTime_summary
```

The next set of results show the number of correct responses grouped by method.

```{r error-rates, echo=FALSE}
#number of correct responses grouped by method
aggregate(my_data$all_q, by = list(my_data$method), FUN = sum)
```

## Rate Correct Score
To analyse both error rates and completion times together, we use the RATE CORRECT SCORE ("A Comparison of Methods to Combine Speed and Accuracy Measures of Performance: a rejoining on the binning procedure" - Andre Vandierendonck). 

RCS  = #correct responses / average response time

and can be interpreted as the number of correct responses per second of task activity. (this value has been square root transformed).
```{r rcs-summary, echo=FALSE}
#Rate Correct Score (rcs) -> number of correct responses per second of activity
rcs_summary <- my_data %>% group_by(method) %>% get_summary_stats(rcs, type = "mean_sd")
rcs_summary
```

## ASSUMPTION TESTING

```{r outlier-test, include=FALSE}
my_data %>% group_by(method) %>% identify_outliers(completion_time);
my_data %>% group_by(method) %>% identify_outliers(rcs);
```

### Normality test using Shapiro-wilk test.

Completion Times
```{r normal-completion-time-test}
my_data %>% group_by(method) %>% shapiro_test(completion_time);
```

RCS
```{r normal-rcs-test}
my_data %>% group_by(method) %>% shapiro_test(rcs);
```

Error Rate is not presented as it is not normal

### Homogeneity of Variances using Levene's test.

Completion Times
```{r variance-completion-time-test}
leveneTest(my_data$completion_time~my_data$method);
```

RCS
```{r variance-rcs-test}
leveneTest(my_data$rcs~my_data$method);
```

### QQ-plots

Completion Time
```{r qq-completiont-time, echo=FALSE}
ggqqplot(my_data,"completion_time", ggtheme = theme_bw()) + 
  facet_grid(~method, labeller = "label_both");
```

RCS
```{r qq-rcs, echo=FALSE}
ggqqplot(my_data,"rcs", ggtheme = theme_bw()) + 
  facet_grid(~method, labeller = "label_both");
```

## ANOVA

Completion Time
```{r ANOVA-completion-time, message=FALSE, warning=FALSE}
ezANOVA(data = my_data, dv = .(completion_time), wid = .(pid), between = .(method), type = 3, detailed = T);
```

RCS
```{r ANOVA-rcs, message=FALSE, warning=FALSE}
ezANOVA(data = my_data, dv = .(rcs), wid = .(pid), between = .(method), type = 3, detailed = T);
```

Removing the outliers from the data does not change the ANOVA results.
```{r outlier-dicard, message=FALSE, warning=FALSE}
outlier_discard <- c(2,11,28,34,39,64)
my_data_outlier_discarded <- my_data[!(my_data$pid %in% outlier_discard),]

ezANOVA(data = my_data_outlier_discarded, dv = .(rcs), wid = .(pid), between = .(method), type = 3, detailed = T);
```

## KRUSKAL-WALLIS
```{r kruskal-rcs, message = FALSE, warning = FALSE}
kruskal.test(rcs~method, data = my_data)
```
## ANCOVA - COMPLETION TIME

### Assumptions

#### Linearity Assumption

Completion Time vs Age
```{r age_CT_linear, echo=FALSE, message=FALSE, warning=FALSE}
ggscatter(
  my_data, x = "age", y = "completion_time",
  facet.by  = "method", 
  short.panel.labs = FALSE
)+
  stat_smooth(method = "loess", span = 0.9)
```

The other covariates (gender, education, mid-air, problem solving) are categorical and plotting their data to assess their linearity does not make sense.

#### Homogeneity of Regression Slopes

According to https://www.datanovia.com/en/lessons/ancova-in-r/, the test for homogeniety of regression slopes is an ANOVA with the covariates coming in first (age, gender, education, mid-air, problem solving) then the grouping variable (method) and then every combination of intearations.

```{r regress_slopes_CT}
my_data %>%
  anova_test(completion_time ~ 
               age + 
               gender + 
               education + 
               mid_air + 
               problem_solving + 
               method +
               age * method +
               gender * method +
               education * method +
               mid_air * method + 
               problem_solving * method +
               age * gender * method +
               age * education * method +
               age * mid_air * method +
               age * problem_solving * method +
               gender * education * method +
               gender * mid_air * method + 
               gender * problem_solving * method +
               education * mid_air * method +
               education * problem_solving * method +
               mid_air * problem_solving * method +
               age * gender * education * method +
               age * gender * mid_air * method +
               age * gender * problem_solving * method +
               age * education * mid_air * method +
               age * education * problem_solving * method +
               age * mid_air * problem_solving * method +
               gender * education * mid_air * method + 
               gender * education * problem_solving * method +
               gender * mid_air * problem_solving * method +
               education * mid_air * problem_solving * method + 
               age * gender * education * mid_air * problem_solving * method)
```

This does not make much sense to me as using (age * gender * education * mid_air * problem_solving * method) should give the same results but Senuri mentioned that order matters which is why the above monstrosity was constructed.

Regression slope for completion time and age seems to deviate from the rest.

#### Normality 

```{r model-ancova-CT, include = FALSE}
model.ct <- lm(completion_time ~ age + gender + education + mid_air + problem_solving + method, data = my_data)

model.ct.metrics <- augment(model.ct) %>%
  select(-.hat, -.sigma, -.fitted, -.se.fit) # Remove details
head(model.ct.metrics, 3)
```

Completion Times normality test using shapiro-wilk test. No significant results
```{r normality-ancova-CT, echo=FALSE}
shapiro_test(model.ct.metrics$.resid)
```

#### Homogeneity of variances
Completion Times Homogeneity of variances test using levene's test. No significant results.
```{r variance-CT, echo=FALSE}
levene_test(.resid ~ method, data = model.ct.metrics)
```

#### Outliers

Completion Time - no outliers were found
```{r outliers-CT, echo=FALSE}
model.ct.metrics %>% 
  filter(abs(.std.resid) > 3) %>%
  as.data.frame()
```

#### Computation

ANCOVA

DV = completion time, IV = method, covariates = age, gender, education, mid_air, problem_solving, method
```{r ancova-CT}
res.aov <- my_data %>% 
  anova_test(completion_time ~ age + gender + education + mid_air + problem_solving + method, type = 2)
get_anova_table(res.aov)
```

## ANCOVA - RCS

### Assumptions

#### Linearity Assumption

RCS vs Age
```{r age_rcs_linear, echo=FALSE, message=FALSE, warning=FALSE}
ggscatter(
  my_data, x = "age", y = "completion_time",
  facet.by  = "method", 
  short.panel.labs = FALSE
)+
  stat_smooth(method = "loess", span = 0.9)
```

The other covariates (gender, education, mid-air, problem solving) are categorical and plotting their data to assess their linearity does not make sense.

#### Homogeneity of Regression Slopes

According to https://www.datanovia.com/en/lessons/ancova-in-r/, the test for homogeniety of regression slopes is an ANOVA with the covariates coming in first (age, gender, education, mid-air, problem solving) then the grouping variable (method) and then every combination of intearations.

```{r regress_slopes_rcs}
my_data %>%
  anova_test(rcs ~ 
               age + 
               gender + 
               education + 
               mid_air + 
               problem_solving + 
               method +
               age * method +
               gender * method +
               education * method +
               mid_air * method + 
               problem_solving * method +
               age * gender * method +
               age * education * method +
               age * mid_air * method +
               age * problem_solving * method +
               gender * education * method +
               gender * mid_air * method + 
               gender * problem_solving * method +
               education * mid_air * method +
               education * problem_solving * method +
               mid_air * problem_solving * method +
               age * gender * education * method +
               age * gender * mid_air * method +
               age * gender * problem_solving * method +
               age * education * mid_air * method +
               age * education * problem_solving * method +
               age * mid_air * problem_solving * method +
               gender * education * mid_air * method + 
               gender * education * problem_solving * method +
               gender * mid_air * problem_solving * method +
               education * mid_air * problem_solving * method + 
               age * gender * education * mid_air * problem_solving * method)
```

This does not make much sense to me as using (age * gender * education * mid_air * problem_solving * method) should give the same results but Senuri mentioned that order matters which is why the above monstrosity was constructed.

#### Normality 

```{r model-ancova-rcs, include = FALSE}
model.rcs <- lm(rcs ~ age + gender + education + mid_air + problem_solving + method, data = my_data)

model.rcs.metrics <- augment(model.rcs) %>%
  select(-.hat, -.sigma, -.fitted, -.se.fit) # Remove details
head(model.rcs.metrics, 3)
```

rcs normality test using shapiro-wilk test. SIGNIFICANT RESULTS
```{r normality-ancova-rcs, echo=FALSE}
shapiro_test(model.rcs.metrics$.resid)
```

#### Homogeneity of variances
rcs homogeneity of variances test using levene's test. SIGNIFICANT RESULTS
```{r variance-rcs, echo=FALSE}
levene_test(.resid ~ method, data = model.rcs.metrics)
```

#### Outliers

rcs - no outliers were found
```{r outliers-rcs, echo=FALSE}
model.rcs.metrics %>% 
  filter(abs(.std.resid) > 3) %>%
  as.data.frame()
```

#### Computation

The RCS Dependent Variable breaks both assumptions of normality and homogeneity of variances when taking covariates into account. ANCOVA may not be appropriate here.

ANCOVA

DV = rcs, IV = method, covariates = age, gender, education, mid_air, problem_solving, method
```{r ancova-rcs}
res.aov <- my_data %>% 
  anova_test(rcs ~ age + gender + education + mid_air + problem_solving + method, type = 2)
get_anova_table(res.aov)
```